---
title: "p8105_hw3_mp3745"
author: "Matthew Parker"
date: "10/08/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(httr)
library(jsonlite)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 1

```{r}
# Download data
library(p8105.datasets)
data("instacart")

instacart = instacart %>%
  janitor::clean_names()
```

The instacart dataset consists of `r nrow(instacart)` observations of `r ncol(instacart)` variables. The data consists of information regarding orders and the items that were included. In total, there are `r instacart %>% pull(order_id) %>% n_distinct` orders recorded in the data.

## Distribution of order sizes
```{r}
instacart %>% 
  group_by(order_id) %>% 
  summarize(
    order_size = n()
  ) %>% 
  ggplot(aes(x = order_size)) +
  geom_histogram() +
  scale_x_continuous(
    breaks = c(0, 5, 10, 15, 20, 30, 40, 60, 80)
  )
```

It looks like the vast majority of orders have less than 20 items.

## Orders by day of week
```{r}
instacart %>% 
  group_by(order_dow) %>% 
  summarize(
    order_count = n_distinct(order_id)
  ) %>%
  mutate(
    order_dow = recode(
      order_dow, 
      "0" = "Sunday", 
      "1" = "Monday", 
      "2" = "Tuesday", 
      "3" = "Wednesday", 
      "4" = "Thursday", 
      "5" = "Friday",
      "6" = "Saturday"
    )
  ) %>% 
  knitr::kable()
```

Unsurprisingly, we can see that most orders from instacart are placed on Sunday.

## Most popular aisles
```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    item_count = n()
  ) %>%
  arrange(desc(item_count)) %>% 
  top_n(10) %>%
  knitr::kable(caption = "Top 10 aisles by number of items ordered")
```

There are `r instacart %>% pull(aisle) %>% n_distinct()` different aisles that items can be ordered from. The fresh vegetables aisle has the most items ordered from it. 


```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    item_count = n()
  ) %>% 
  filter(item_count > 10000) %>%
  ggplot(aes(x = reorder(aisle, item_count), y = item_count)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  xlab("aisle")
```

This plot shows the number of items ordered per aisle for the aisles with over 10,000 items ordered. Again, we can see that the fresh vegetables aisle had the most items ordered.


```{r}
instacart %>% 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>%
  summarize(
    item_count = n()
  ) %>% 
  ungroup() %>%
  group_by(aisle) %>%
  mutate(item_count_ranking = min_rank(desc(item_count))) %>%
  filter(item_count_ranking <= 3) %>% 
  arrange(aisle, item_count_ranking) %>% 
  select(-item_count_ranking) %>% 
  knitr::kable(caption = "Top 3 items ordered from baking ingredients, dog food car, and packaged vegetables fruits aisles")
```

From this table, we can see that light brown sugar, organix grain free chicken & vegetable dog food, and organic baby spinach were the most popular items from the baking ingredients, dog food care, and packaged vegetables fruits aisles, respectively.


```{r}
instacart %>% 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>% 
  summarize(
    mean_order_hour = round(mean(order_hour_of_day), 0)
  ) %>% 
  mutate(
    order_dow = recode(
      order_dow, 
      "0" = "Sunday", 
      "1" = "Monday", 
      "2" = "Tuesday", 
      "3" = "Wednesday", 
      "4" = "Thursday", 
      "5" = "Friday",
      "6" = "Saturday"
    )
  ) %>% 
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_order_hour"
  ) %>% 
  knitr::kable(caption = "Mean hour of day that Coffee Ice Cream and Pink Lady Apples are ordered by day of week")
```

From the table we can see that pink lady apples tend to be ordered earlier in the day compared to coffee ice cream, except for Friday and Saturday.


# Problem 2

```{r, cache=TRUE}
# Download data
var_names = GET("https://chronicdata.cdc.gov/views/acme-vg9e.json") %>%
  content("text") %>%
  fromJSON() %>% 
  .[["columns"]] %>% 
  .[["name"]] %>% 
  .[-23]

brfss_smart2010 = 
  GET("https://chronicdata.cdc.gov/views/acme-vg9e/rows.json") %>% 
  content("text") %>%
  fromJSON() %>% 
  .[["data"]]

row_as_tibble = function(row_as_list, var_names) {
  var_list = row_as_list[9:32]
  names(var_list) = var_names 
  var_list[sapply(var_list, is.null)] <- NULL
  as_tibble(var_list, validate = FALSE)
}

brfss_smart2010 = 
  brfss_smart2010 %>% 
  map(.x = ., ~row_as_tibble(.x, var_names)) %>% 
  bind_rows
```

```{r}
# data cleaning
brfss_smart2010_ovr_hlth = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  filter(response == c("Excellent", "Very good", "Good", "Fair", "Poor")) %>% 
  mutate(
    response = fct_relevel(as_factor(response), "Poor", "Fair", "Good", "Very good", "Excellent")
  ) %>%
  arrange(response)
```

States observed in 2002 at 7 or more locations
```{r}
brfss_smart2010_ovr_hlth %>% 
  filter(year == 2002) %>% 
  separate(locationdesc, into = c("State", "County"), sep = "-") %>% 
  group_by(State) %>% 
  summarize(
    num_locations = n_distinct(County)
  ) %>% 
  filter(num_locations >= 7) %>% 
  knitr::kable()
```

There were 6 states observed at 7 or more locations in 2002.


States observed in 2010 at 7 or more locations
```{r}
brfss_smart2010_ovr_hlth %>% 
  filter(year == 2010) %>% 
  separate(locationdesc, into = c("State", "County"), sep = "-") %>% 
  group_by(State) %>% 
  summarize(
    num_locations = n_distinct(County)
  ) %>% 
  filter(num_locations >= 7) %>% 
  knitr::kable()
```

There were 14 states observed at 7 or more locations in 2010, which is more than 2002.


```{r}
brfss_smart2010_ovr_hlth %>% 
  filter(response == "Excellent") %>% 
  group_by(year, locationabbr) %>% 
  summarize(
    avg_data_value = mean(as.numeric(data_value), na.rm = TRUE)
  ) %>% 
  select(year, locationabbr, avg_data_value) %>% 
  ungroup() %>% 
  ggplot(aes(x = year, y = avg_data_value, group = locationabbr, color = locationabbr)) + 
  geom_line()
```

From this plot, we can see an overall deceasing trend in the average data value each year at the state level from 2002 to 2010. For the most part, the states are fairly clustered together, with some having a higher average data value than others.


Looking at data value in NY by response for 2006 and 2010.
```{r}
brfss_smart2010_ovr_hlth %>% 
  filter(year == c("2006", "2010"), locationabbr == "NY") %>%
  mutate(data_value = as.numeric(data_value)) %>% 
  ggplot(aes(x = response, y = data_value, color = response)) +
  geom_violin() +
  facet_grid(~year)
```

The distribution of data value in NY across responses for 2006 and 2010 look relatively similar, however, the distribution for excellent response has got narrower and decreased in 2010, whereas for fair response, it has spread apart.


# Problem 3

```{r}
# load data
accel_data = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_count"
  ) %>% 
  mutate(
    minute = as.numeric(minute),
    day = fct_relevel(as_factor(day), "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"),
    is_weekend = if_else(
      day == "Saturday", TRUE, if_else(
        day == "Sunday", TRUE, FALSE
      )
    )
  )
```

In this dataset, we have five weeks of accelerometer data collect on a 63 year-old male with BMI 25. The tidied dataset has the following variables: `r colnames(accel_data)`. There are a total of `r nrow(accel_data)` rows.

```{r}
accel_data %>% 
  group_by(week, day, is_weekend) %>% 
  summarize(
    tot_daily_activity = sum(activity_count)
  ) %>% 
  select(day, tot_daily_activity, is_weekend) %>% 
  knitr::kable()
```

Based on the above table looking at total daily activity, it is hard to tell if there are any trends. Towards the end of the five week period, it seems total daily activity tends to be lower around the weeknd though.


```{r}
accel_data %>% 
  ggplot(aes(x = minute, y = activity_count, color = day)) +
  geom_line()
```

From this plot, it seems that this man is more active earlier in the day on the weekend, especially on Sunday. Conversely, he is more active later in the day during the week days.